---
title: "Plotting Marker Locations in Genome"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv",
    chr_len = "inputs/Otsh_v2.0_chrom_lengths.tsv"
  )
  # outputs:
  output_list <- list(
    pdf = "results/genomic-locations-plot/gen-loc.pdf",
    pdfcrop = "results/genomic-locations-plot/gen-loc-crop.pdf",
    pdftex = "tex/images/genomic-locations-plot.pdf"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```


# Introduction

We desire a simple plot for the supplement showing us the location of different markers,
colored according to their category.  Should be pretty easy.  We will use the Otsh_v2.0
coordinates.  

```{r}
library(tidyverse)
library(ggchicklet)  # for geom_rrect().  Get it with remotes::install_github("hrbrmstr/ggchicklet")

amplicons <- read_csv(input_list$locus_info) %>%
  mutate(Type = LocusOrigin, .before = AmpliconName)

chroms <- read_tsv(input_list$chr_len) %>%
  filter(chrom != "MT") %>%
  mutate(chrom_dbl = as.numeric(chrom))
```


First, we want to plot the chromosomes as rounded rectangles.

```{r}
chroms2 <- chroms %>%
  mutate(
    chrom_y = 35 - chrom_dbl,
    mb = num_bases / 1e6
  )

hh <- 0.35  # the half-height of the chromosome
  
base <- ggplot() +
  geom_rrect(
    data = chroms2, 
    mapping = aes(ymin = chrom_y - hh, ymax = chrom_y + hh, xmin = 0, xmax = mb),
    fill = NA,
    colour = "gray"
    ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  ) +
  scale_x_continuous(breaks = seq(10, 100, by = 10)) +
  scale_y_continuous(breaks = 1:34, labels = 34:1)
```

Now, we get the markers on there
```{r}
amplicons2 <- amplicons %>%
  filter(!str_detect(Otsh_v2.0_chromosome, "NW")) %>%
  mutate(
    chrom_y = 35 - as.numeric(Otsh_v2.0_chromosome),
    mb = AmpliconStartPos / 1e6,
    Type = factor(Type, levels = c("SNPtype", "Microhap", "RoSA", "LFAR", "WRAP", "VGLL3", "Six6", "Coho"))
  )

final <- base +
  geom_segment(
    data = amplicons2,
    mapping = aes(x = mb, xend = mb, y = chrom_y - hh, yend = chrom_y + hh, colour = Type),
    lineend = "butt"
  ) +
  scale_color_brewer(type = "qual", palette = "Set1") +
  xlab("Genomic coordinate (in megabases)") +
  ylab("Chromsome (in Otsh_v2.0)")

final
```


Write the thing out to a PDF file.
```{r}
ggsave(final, file = output_list$pdf, width = 7, height = 8.5)
```

Then pdfcrop it and put it where we can include it in the Supplement.
```{r}
# let's crop that down for use, too
CALL <- paste("pdfcrop ", output_list$pdf, collapse = " ")
system(CALL)

file.copy(from = output_list$pdfcrop, to = output_list$pdftex, overwrite = TRUE)
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```


